---
title: "Lab2"
author: "Jamil Ahmed"
date: "2024-01-30"
output: 
  html_document:
    toc: True
    toc_float: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

```{r}
# This is the directory structure to my data

getwd()
```

# Task 2

```{r}

#	Read the file and obtain the first six lines of the data using “head()”

dird="C:/Users/Ahmed/OneDrive/Desktop/Math Labs/Lab 2/DATAxls/"
#function to read data

myread=function(csv){
  fl=paste(dird,csv,sep="")
  read.table(fl,header=TRUE,sep=",")
}
mpg.df=myread("EPAGAS.csv")
head(mpg.df)
```

# Task 3

```{r}
# Make the object mpg, the number of miles per gallon vector.
mpg= mpg.df$MPG
head(mpg)
```

```{r}
# Transform the mpg variable to z and verify these results.
z=(mpg-mean(mpg))/sd(mpg)
class(z)
```

```{r}
# Using z, find the values of mpg that are possible outliers.

z[abs(z)>=2 & abs(z)<=3]
```

```{r}
# Using z, find the values of mpg that defined as outliers.
mpg[abs(z)>3]
``` 
 
```{r}
# Plot the outliers in red and the possible outliers in blue
mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))  

library(lattice)
dotplot(mpg,col=mycol, main="Plot of MPG")
```


# Task 4

```{r}
#	Make a boxplot of the mpg variable
boxplot(mpg,main="Boxplot of MPG",xlab= "MPG",col="Black",horizontal=TRUE,notch=TRUE)
```

Predict the proportion of data within 2 standard deviations of the mean of the mpg data.

According to Chebyshev:

Within 2 standard deviations of the mean (k=2)
At least 1-1/4=3/4
3/4

```{r}
 # Use R to calculate the exact proportion within 2 standard deviation of the mean.
length(mpg[abs(z)<2])/length(mpg)
```
Does Chebyshev agree with the data?

No, Chebyshev's says that 3/4 or 75% of the data should be within 2 standard deviations of the mean. However, when we used R, we found that proprotion to be .96, or 96%


Now use the empirical rule, what proportion (according to the rule) of the data should be within 2 standard deviations of the mean?

According to the Empirical rule, 95% of the data should be within 2 standard devations

How well does it correspond?
It corresponds pretty well, Using R, we saw that 96% of the data is within 2 standard deviations of the mean, Where the Empirical rule estimates that 95% of the data is within 2 standard deviations.


Is the Empirical rule valid in this case? Why?
Yes it is valid. The Empirical rule is an approximation. With a large enough sample size we should get closer and closer to 95%. We got 96%, but that does not invalidate the Empirical rule.

Here is a Histogram visualizing the data - We can see that the majority of the data is within 2 standard deviations of the mean (depicted by the green lines) further validating the Empirical rule.
```{r}

hist(mpg, breaks = 20, main = "Histogram", xlab = "MPG", ylab = "Frequency")

# Add lines for mean and standard deviations
abline(v = mean(mpg), col = "red", lwd = 2)  # Mean
abline(v = mean(mpg) + c(-1, 1) * sd(mpg), col = "blue", lty = 2)  # ±1 SD
abline(v = mean(mpg) + c(-2, 2) * sd(mpg), col = "green", lty = 2)  # ±2 SD
abline(v = mean(mpg) + c(-3, 3) * sd(mpg), col = "orange", lty = 2)  # ±3 SD
```

