---
title: "What the package does"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{What the package does}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MATH4753AHMED2024)
```

# Introduction
These are several functions created or given throughout the course of MATH 4753 taught by Dr. Wayne Stewart. This documen will have 4 functions that will be used and have an explanation afterwords describing the function and what it does.

I will be using the FIREDAM data set, it has been implemented into the package as "ddt"

# 1. zscore

```{r}
zscore(ddt$DISTANCE)
```

The Z score measures how many standard deviations a data point is from the mean of the vector.
It's commonly used in outlier detection to identify observations that deviate significantly from the mean of the dataset.

each data point is calculated and then the zscore for the data point is put out in order of which the vector was put in.

# 2. mybin
```{r}
MATH4753AHMED2024::mybin(iter=100,n=10,p=0.5)
```
The `mybin` function is designed to simulate a binomial distribution. It takes three parameters as input: `iter`, `n`, and `p`. The `iter` parameter determines the number of iterations to perform, with each iteration representing a sampling from the binomial distribution. The `n` parameter specifies the number of bins in the simulation, while `p` denotes the probability of success for each trial in the binomial distribution. The function then computes the number of successes in each iteration and generates a barplot representing the distribution of these successes. Finally, it returns a numeric vector containing the proportions of successes for each possible outcome in the simulation, normalized to sum to 1.

# 3. myquad

```{r}
MATH4753AHMED2024::myquad(ddt$DISTANCE)
```
Quadratic Function  This function calculates the quadratic function for a given numeric vector.
from the beginning of class

# 4 ntickets

```{r}
MATH4753AHMED2024::ntickets(150,0.1,0.05)
```

The `ntickets` function is designed to calculate the number of tickets to be sold for a flight, taking into account factors such as the number of seats available, the probability of a "show" (the likelihood that passengers will actually board the flight), and the probability of overbooking. It offers two calculation methods: one based on a discrete distribution and the other using a normal approximation. 

The function calculates the number of tickets to be sold (`nd`) using the discrete distribution method and the number of tickets (`nc`) using the normal approximation. It then prints out a named list containing `nd`, `nc`, `N` (number of seats), `p` (probability of a "show"), and `gamma` (probability of overbooking). Additionally, it generates plots of the objective function versus `n` for both the discrete and continuous cases, providing visual insights into the optimization process.


# data set

```{r}
MATH4753AHMED2024::ddt
```

# devcheck last 10 lines
✔  checking DESCRIPTION meta-information ... 
✔  checking top-level files
✔  checking for left-over files ...
✔  checking index information ... 
✔  checking package subdirectories ... 
✔  checking R files for non-ASCII characters ... 
✔  checking R files for syntax errors ... 
✔  checking whether the package can be loaded ... 
✔  checking whether the package can be loaded with stated dependencies ... 
✔  checking whether the package can be unloaded cleanly ... 
✔  checking whether the namespace can be loaded with stated dependencies ... 
✔  checking whether the namespace can be unloaded cleanly (344ms)
✔  checking loading without being on the library search path ... 
✔  checking dependencies in R code (582ms)
✔  checking S3 generic/method consistency ... 
✔  checking replacement functions ... 
✔  checking foreign function calls ... 
✔  checking R code for possible problems (2.8s)
✔  checking Rd files ... 
✔  checking Rd metadata ... 
✔  checking Rd line widths ... 
✔  checking Rd cross-references ... 
✔  checking for missing documentation entries ... 
✔  checking for code/documentation mismatches (692ms)
✔  checking Rd \usage sections (583ms)
✔  checking Rd contents ... 
✔  checking for unstated dependencies in examples ... 
✔  checking contents of 'data' directory ...
✔  checking data for non-ASCII characters ... 
✔  checking LazyData ...
✔  checking data for ASCII and uncompressed saves ... 
✔  checking installed files from 'inst/doc'
✔  checking files in 'vignettes'
✔  checking examples (573ms)
✔  checking for unstated dependencies in vignettes ... 
✔  checking package vignettes in 'inst/doc' ...
✔  checking re-building of vignette outputs (1.6s)
✔  checking for non-standard things in the check directory
✔  checking for detritus in the temp directory ...
   
   
── R CMD check results ───────────────────────── MATH4753AHMED2024 0.1.0 ────
Duration: 19.2s

0 errors ✔ | 0 warnings ✔ | 0 notes ✔

R CMD check succeeded
